\chapter{Algorithm Design}\label{f9cd972}

\subsection{Introduction}\label{dd8a193}

\def\NP{\mathsf{NP}}
\def\P{\mathsf{P}}
\def\coNP{\mathsf{coNP}}
\def\SAT{\mathsf{SAT}}
\def\VC{\mathsf{VC}}
\def\CONT{\mathsf{CONT}}
\def\Yes{\text{Yes}}
\def\No{\text{No}}
\def\binrep{\{0,1\}^*}

\Definition{Vertex cover}\label{b530c81}

Let $G=(V,E)$ be an undirected graph. A vertex cover $U\subseteq V$ satisfies
$$
  (u,v)\in E\implies u\in U\lor u\in U.
$$

In other words, every edge in $E$ has at least one endpoint in the vertex cover
$U$. Such a set is said to \textit{cover} the edges of $G$.

\Definition{Flow network}\label{cc18bdc}

A \textit{flow network} is a directed graph $G=(V,E)$ with a single
\textit{source node} $s$ and a single \textit{target node} $t$, as well as a
positive number $c(e)$ for each edge $e\in E$, called the capacity of $e$.

\Definition{Flow}\label{b06cb3d}

Let $G$ be a flow network. A \textit{flow} on $G$ is given by a positive number
$f(e)$ for each edge $e$ in $G$ satisfying the following two constraints:
\begin{itemize}
  \item \textbf{Capacity constraints.} For each edge $e\in E$, we have
        $0\leq f(e)\leq c(e)$
  \item \textbf{Flow conservation.} For each vertex $v\in V$ that is
        not the source or target vertex,
\end{itemize}
$$
  \sum_{e\text{ into }v}f(e)=\sum_{e\text{ leaving }v}f(e)
$$

The \textit{value} of flow $f$ is all of the flow leaving $s$:
$$
  \text{val}(f):=\sum_{e\text{ leaving }s}f(e)
$$

where $s$ is the source node of $G$.

\Problem{Max Flow}\label{e67a151}

\textbf{Input:} A flow network $G$ with source $s$, target $t$, and positive
edge capacities $c(e)$ for $e\in E$.

\textbf{Output:} A flow $f$ with the maximum value.

\Definition{Residual graph}\label{b5b85fc}

Let $G$ be a flow network and let $f$ be a flow on $G$. The \textit{residual
graph} of $G$ and $f$, denoted by $G_f$, is the directed graph defined as
follows:

The vertices of $G_f$ are the same as the vertices of $G$.

For each edge $e=(u,v)$ in $G$, if $f(e)<c(e)$ then we add the edge $(u,v)$ to
$G_f$, labelled with the number $c(e)-f(e)$. If $f(e)>0$, then we also add the
edge $(v,u)$ to $G_f$, labelled with the number $f(e)$.

All paths from $s$ to $t$ in the residual graph correspond to a sequence where
flow can be re-routed to increase its value.

\Definition{Binary representation}\label{c043951}

Let
$$
  \binrep:=\{\epsilon,0,1,00,01,10,11,100,101,\ldots\}
$$

be the set of all finite binary strings. (where $\epsilon$ is the empty
string.)

\Definition{Decision problem}\label{c2ad5af}

A decision problem $L$ is a subset of $\binrep$. The computational task
corresponding to $L$ is ``Given a string $x\in\binrep$, is $x\in L$?"

\Problem{$L$-membership problem}\label{acff8f4}

\textbf{Input:} A boolean string $x$.

\textbf{Output:} Decide if $x\in L$.

\Example{Rewriting problems as $L$-membership problems}\label{d7cf97b}

\begin{itemize}
  \def\B{\binrep}
  \item\textit{Graph Connectivity.} Given a graph $G=(V,E)$, is it
        connected?
        \begin{align*}
          L &=\{x\in\B\mid x\text{ encodes a connected graph}\} \\
            &=\{G \mid G \text{ is a connected graph}\}
        \end{align*}
  \item\textit{Max Flow (Decision Version).} Given a flow network
        $G$ and a postive integer $k$, does the max flow on $G$ have
        value $\geq k$?
        \begin{align*}
          L &=\{x\in\B\mid x\text{ encodes a $(G,k)$ such that $\text{val}(G)\geq k$}\} \\
            &=\{(G,k)\mid\text{val}(G)\geq k \}
        \end{align*}
  \item\textit{Sum.} Given $a,b,c\in\Z$, does $a+b=c$?
        \begin{align*}
          L &=\{x\in\B\mid x\text{ encodes $(a,b,c)$ such that $a+b=c$}\} \\
            &=\{(a,b,c)\mid a+b=c\}
        \end{align*}
\end{itemize}

Choice of encoding is important and definitely affects runtime. However, for
discussion we will assume that the most natural and succint encoding is chosen.

\Definition{Polynomial-time algorithms}\label{ee3be28}

An algorithm $A$ runs in \textit{polynomial time} if $\exists c\in\R$ s.t.
$\forall x\in\binrep$, $A$ terminates after $O(|x|^c)$ computation steps.

A decision problem $L$ is \textit{polynomial-time computable} if there exists a
polynomial-time $A$ s.t. $\forall x\in\binrep$, $x\in L\iff A(x)=\Yes$.

We define
$$
  \P:=\{L\subseteq\binrep\mid L\text{ is polynomial-time computable}\}
$$

The complexity class $\P$ is our proxy for \textit{efficiently computable
languages}. (``language" is another way to refer to $L$, in addition to
``problem".)

\Definition{Vertex Cover Problem}\label{bd1de5f}

$$
  \VC:=\{(G,k)\mid G\text{ is a graph with a
    \href{b530c81}{vertex cover} of size}\leq k\}
$$

\Definition{Satisfiability Problem}\label{d6893fe}

$$
  \mathsf{FORMSAT}:=\{F\mid F\text{ is a satisfiable boolean formula}\}
$$

\Definition{Nondeterministic polynomial-time algorithms}\label{d93050e}

A decision problem $L$ has a polynomial-time verifier if there is a polynomial
time algorithm $B$ taking two strings $x,y$ as input, and a polynomial $p(n)$
such that
$$
  x\in L\iff\exists y\in\binrep,|y|\leq p(|x|):B(x,y)=\Yes
$$

The complexity class
$$
  \mathsf{NP}:=\{L\subseteq\binrep\mid L\text{ has a polynomial-time verifier}\}
$$

\begin{itemize}
  \item A polynomial-time verifier for $\VC$ would take a graph $(G,k)$ and a
        proposed vertex cover $U$ and check if $|U|\leq k$ and that $U$ is a
        vertex cover.
  \item A polynomial-time verifier for $\SAT$ would take in a boolean formula
        $F$ and a proposed assignment $x$ and check if $F(x)=\Yes$.
\end{itemize}

Observe that $\P\subseteq\NP$. If $L$ has a polynomial-time algorithm, then it
also has a polynomial-time verifier.

\begin{proof}
  If $L\in\P$, then \href{ee3be28}{by definition} there exists a polynomial-time algorithm $A$ with
  $$
    x\in L\iff A(x)=\Yes
  $$

  Then, following the \href{d93050e}{defintion of $\NP$}, we need to find a
  polynomial-time verifier $B$ such that
  $$
    x\in L\iff\exists y\in\binrep,|y|\leq p(|x|):B(x,y)=\Yes
  $$

  But we can simply use $B(x,y):=A(x)$.
\end{proof}

\Definition{Complement of a decision problem}\label{f636043}

The complement of a decision problem $L$ is defined as
$$
  \overline L=\binrep\setminus L:=\{x\in\binrep\mid x\notin L\}
$$

Note that $\binrep=L\cup\overline L$ for any decision problem $L$.

\Exercise{}\label{da1567b}

Prove that if $L\in\P$ then $\overline L\in\P$.

\begin{proof}
  If $L\in\P$, then \href{ee3be28}{by definition} there exists a
  polynomial-time algorithm $A$ with
  $$
    x\in L\iff A(x)=\Yes
  $$

  Then, following the \href{ee3be28}{defintion of $\P$}, we need to find a
  polynomial-time algorithm $B$ such that
  $$
    x\in\overline L\iff B(x)=\Yes
  $$

  But we can simply use $B(x):=\neg A(x)$.
\end{proof}

\Definition{$\mathsf{coNP}$}\label{a92ed96}

The complexity class $\mathsf{coNP}$ is defined as
$$
  \mathsf{coNP}:=\{L\mid\overline L\in\mathsf{NP}\}
$$

For example, recall that $\mathsf{\href{d6893fe}{SAT}}=\{F\mid F\text{ is a
satisfiable boolean formula}\}$. Then
$$
  \overline{\SAT}=\left\{x\in\binrep\, \middle\vert
  \begin{array}{l}
    \text{$x$ is an invalid encoding of a formula, or} \\
    \text{$x$ encodes an unsatisfiable boolean formula}
  \end{array}
  \right\}
$$

But given $x\in\binrep$, it is easy to test its validity as a boolean formula,
hence we focus on the second constraint:
$$
  \CONT:=\{F\mid F\text{ is an unsatisfiable boolean formula}\}
$$

Note that since $\overline{\CONT}=\SAT\in\NP$, we have that
$\overline{\CONT}\in\NP$.

However, is $\CONT\in\NP$? Observe that $F\in\CONT$ if and only if \textit{for
every} assignment $x$ to the variables of $F$, we have that $F(x)=\No$. Since
there are $2^n$ assigments to check, it is not clear how to encode this
checking procedure into a single polynomial-sized certificate. For this reason,
many researchers conjecture that $\NP\neq\coNP$.

\Remark{The complexity class $\mathsf{NP}\cap\mathsf{coNP}$}\label{e6ea03d}

We can show that $\P\subseteq\coNP$ by
$$
  L\in\P\implies\overline L\in\P\implies\overline L\in\NP\implies L\in\coNP
$$

And since $\P\subseteq\NP$, we have
$$
  \P\subseteq\NP\cap\coNP
$$

% Sunday, Nov 12 2:01 AM
% I have a sleep cycle problem

% It is not known how to convert CNF into DNF efficiently (P-time).