\chapter{Statistics I: Examples}\label{fc5d59d}

\subsection{Counting}\label{bb61661}

\Example{}\label{dea0f9c}

A king comes from a family of two children. What is the probability that the
other child is his sister?

\begin{solution}
  The problem is formally re-stated as such: given that one child among two is a
  boy, what is the probability that the other child is a girl?

  The reduced sample space is now just
  $$
    \{(BB), (BG), (GB)\}
  $$

  and hence the answer is $2/3$.
\end{solution}

\Example{}\label{a2a8531}

A family has two children. What is the probability that the younger child is a
girl?

\begin{solution}
  This is equivalent to asking what's the probability that one particular child
  is a girl. To which the answer is $1/2$.
\end{solution}

\Example{}\label{c9b4b6d}

A card is drawn one at a time, without replacement from a deck of 52 playing
cards.

\begin{enumerati}
  \item What is the probability that the first card drawn is an ace?
  \item Given that the first card drawn is an ace, what is the probability that
        the second card drawn is also an ace?
\end{enumerati}

\begin{solution}
  For (i), this is equivalent to asking what's the probability that a particular
  card is an ace. The answer to which is simply $4/52$.

  For (ii), we consider a reduced sample space in which we lost one of the aces
  to the first draw. The solution is hence $3/51$.
\end{solution}

\Example{}\label{a85c9ad}

A total of $8$ balls are sequentially and randomly chosen, without replacement,
from an urn containing $4$ red balls and $6$ blue balls. Given that $5$ of the
$8$ balls are blue, what is the (conditional) probability that the first ball
chosen is blue?

\begin{solution}
  We consider a reduced sample space. Observe that the condition is that $5$ of
  the $8$ chosen balls are blue.

  The problem is then equivalent to asking what's the probability that a
  particular ball among the $8$ is blue.

  The answer is simply $5/8$.
\end{solution}