\chapter{\texttt{core::set\_theory}}\label{f9174c1}

\begin{toc}
  \citem{e7d2b65} % Logic
  \citem{a59f15a} % Basics
\end{toc}

\subsection{Logic}\label{e7d2b65}

\Lemma{Conjuction distributes over disjunction}\label{ea305af}

Let $x,y,z$ be boolean variables. Then
$$
  x\lor(y\land z)=(x\land y)\lor(x\lor z)
$$

\begin{proof}
  This is directly verifiable by listing all 8 possibilities.
\end{proof}

\Lemma{Disjunction distributes over conjunction}\label{ddc349a}

Let $x,y,z$ be boolean variables. Then
$$
  x\land(y\lor z)=(x\lor y)\land(x\land z)
$$

\begin{proof}
  This is directly verifiable by listing all 8 possibilities.
\end{proof}

\subsection{Basics}\label{a59f15a}

\Lemma{Intersection distributes over union}\label{b34061c}

Let $A,B,C$ be sets. Then
$$
  A\cup(B\cap C)=(A\cap B)\cup(A\cup C)
$$

\begin{proof}
  Let $t$ be an element. Let $x:=(t\in A),\ y:=(t\in B),\ z:=(t\in C)$. Then we
  have $t\in A\cup(B\cap C)\iff x\lor(y\land z)\href{ea305af}{\iff}(x\land
  y)\lor(x\lor z)\iff t\in(A\cap B)\cup(A\cup C)$.
\end{proof}

\Lemma{Union distributes over intersection}\label{ec55f1d}

Let $A,B,C$ be sets. Then
$$
  A\cap(B\cup C)=(A\cup B)\cap(A\cap C)
$$

\begin{proof}
  Left as an exercise. Hint: refer to the proof of \autoref{b34061c} and use
  \autoref{ddc349a} instead.
\end{proof}

\Lemma{Cartesian product is distributive over intersection}\label{f05f9b8}

We have that $\times$ (Cartesian product) is distributive over $\cap$. That is
for all sets $A,B,C$,
$$
  (A\cap B)\times C=(A\times C)\cap(B\times C)
$$

Note: $\cap$ is \textbf{not} distributive over $\times$.

\begin{proof}
  The following statements are all equivalent:
  \begin{gather*}
    (x,y)\in(A\cap B)\times C                       \\
    (x\in A\cap B)\land(y\in C)                     \\
    (x\in A)\land(y\in C)\land(x\in B)\land(y\in C) \\
    (x,y)\in A\times C\land(x,y)\in B\times C       \\
    (x,y)\in A\times C\cap B\times C
  \end{gather*}
\end{proof}

\Lemma{Cartesian product of intersections is intersection of a Cartesian products}\label{e2948d3}

Let $U_1,U_2,V_1,V_2$ be sets. Then
$$
  (\textcolor{blue}{U_1}\times\textcolor{red}{V_1})\cap(\textcolor{blue}{U_2}\times\textcolor{red}{V_2})
  =(\textcolor{blue}{U_1}\cap\textcolor{blue}{U_2})\times(\textcolor{red}{V_1}\cap\textcolor{red}{V_2})
$$

\begin{proof}
  The following statements are all equivalent:
  \begin{gather*}
    (u,v)\in(U_1\times V_1)\cap(U_2\times V_2)                                \\
    \bigl((u,v)\in U_1\times V_1\bigr)\land\bigl((u,v)\in U_2\times V_2\bigr) \\
    (u\in U_1)\land(v\in V_1)\land(u\in U_2)\land(v\in V_2)                   \\
    (u\in U_1\cap U_2)\land(v\in V_1\cap V_2)                                 \\
    (u,v)\in(U_1\cap U_2)\times(V_1\cap V_2)
  \end{gather*}
\end{proof}
