\subsection{Isomorphic vector spaces}\label{ef39dd0}

\Proposition{Dimension shows if vector spaces are isomorphic}\label{c10bb88}

Two finite-dimensional vector spaces over $\F$ are isomorphic if and only if
they have the same dimension.

\begin{proof}
  Let $V,W$ be finite-dimensional vector spaces.

  First, suppose that $V$ and $W$ are \href{d0ad6cb}{isomorphic vector spaces}.
  Then there exists an isomorphism $\phi:V\to W$. Since $\phi$ is invertible,
  we have (by \href{bd0d827}{this}, \href{f68db52}{this}, and
  \href{a41ddec}{this}) $\ker\phi=\{0\}$ and $\range\phi=W$. Thus we have both
  $$
    \Null\phi=0,\quad \Rank\phi=\dim W
  $$

  which, when substituted into the \href{e83dffc}{Rank-Nullity Theorem}, gives
  $\dim V=\dim W$.

  Next, suppose $\dim V=\dim W$. Then there exists \href{db2477b}{bases}
  $\iter{v_1}{v_n}$ of $V$ and $\iter{w_1}{w_n}$ of $W$. Let $\phi\in\L(V,W)$
  be defined by
  $$
    \phi(c_1v_1+\ldots+c_nv_n):=c_1w_1+\ldots+c_nw_n
  $$

  Then $\phi$ is a well-defined linear map because $\iter{v_1}{v_n}$ is a basis
  of $V$, and so it maps every element of $v$ unambiguously to one element of
  $W$. Also, $\phi$ is surjective because $\iter{w_1}{w_n}$
  \href{ac574be}{spans} $W$. Furthermore, $\ker\phi=\{0\}$ because of the
  \href{c133a44}{linear independence} of $\iter{w_1}{w_n}$, \href{f68db52}{and
  so} $\phi$ is injective.

  Since $\phi$ is a bijective linear map, by definition it is a vector space
  isomorphism (from $V$ to $W$), and hence $V$ and $W$ are isomorphic.
\end{proof}

\Proposition{$\mathcal L(V,W)$ and $\mathcal F^{m\times n}$ are isomorphic}\label{c172aff}

\texttt{\href{d76dfe6}{use notation};} Let $V,W$ be finite-dimensional vector
spaces, and let $n:=\href{cd4284b}{\dim V}$ and $m:=\dim W$. Then
\href{c70dad0}{$\mathcal M$} is an \href{d0ad6cb}{isomorphism} between $\L(V,W)$
and $\F^{m\times n}$.

\begin{proof}
  By \autoref{abcf830} and \autoref{f95a174}, we have seem that $\mathcal M$ is
  linear. It remains to show that $\mathcal M$ is \href{d205f32}{bijective}.

  Let $\iter{v_1}{v_n}$ be a basis of $V$ and let $\iter{v_1}{v_m}$ be a basis
  of $W$.

  Starting with injectivity, if $T\in\L(V,W)$ and $\mathcal M(T)=0$,
  \href{c70dad0}{then} $T(v_k)=0$ for all $k=\iter1n$. Now because
  $\iter{v_1}{v_n}$ is a basis of $V$, by the \href{d7d1925}{linearity} of $T$
  this implies that $T=0$ (that $T$ is the linear map that sends all vectors to
  $0\in W$). Thus $\mathcal M$ is injective.

  Next, for surjectivity, suppose $A\in\F^{m\times n}$. We can define a $T:V\to
  W$ such that
  \begin{equation*}
    T(v_k)=\sum_{j=1}^m A_{j,k}w_j\with{(k=\iter1n)}\Tag{*}
  \end{equation*}

  From the proof of the \href{f1c27fd}{linear map lemma}, we can infer that $T$
  is well-defined as a function from $V$ to $W$, and that it is linear. Hence
  $T\in\L(V,W)$. Since $\mathcal M(T)=A$ (comparing $(*)$ against
  \href{c70dad0}{definition}), what we've just done is to show that any element
  in $A\in\F^{m\times n}$ has a pre-image in $T\in\L(V,W)$ under $\mathcal M$,
  and hence $\mathcal M$ is surjective.
\end{proof}

\Proposition{$\dim\mathcal L(V,W)=(\dim V)(\dim W)$}\label{eb3bad0}

Let $V,W$ be finite-dimensional vector spaces. Then $\L(V,W)$ is
finite-dimensional and
$$
  \dim\L(V,W)=(\dim V)(\dim W)
$$

\begin{proof}
  This follows from \autoref{f2db1b2}, \autoref{c10bb88}, and \autoref{c172aff}.
\end{proof}
